---
title: "p8105_hw6_mt3866"
output: github_document
date: "2025-11-17"
---

Load libraries
```{r}
library(tidyverse)
library(broom)
library(modelr)
```

### Problem 1
Import data, and create city_state column, exclude 4 cities, only include White & Black, ensure age is numeric:
```{r}
  homicides = read_csv("homicide-data.csv")|>
  janitor::clean_names()|>
  mutate(city_state = str_c(city, state, sep = ","),
    solved = if_else(disposition == "Closed by arrest", 1, 0),
        victim_age = as.numeric(victim_age)
  ) |>
  filter(!city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL"))|>
  filter(victim_race %in% c("White", "Black"))
```

Use glm to fit logistic regression, obtain adjusted OR:
```{r}
baltimore_lr = homicides |>
  filter(city_state == "Baltimore,MD")

balt_model = glm(
  solved ~ victim_age + victim_sex + victim_race,
  data = baltimore_lr,
  family = binomial
)

balt_tidy = tidy(balt_model, exponentiate = TRUE, conf.int = TRUE)
balt_tidy

balt_or_for_male <- balt_tidy |>
  filter(term == "victim_sexMale")

balt_or_for_male
```

Model for all cities, extract adjusted OR:
```{r}
city_models = homicides |>
  group_by(city_state) |>
  nest() |>
  mutate(
    model = map(data, ~ glm(
      solved ~ victim_age + victim_sex + victim_race,
      data = .x,
      family = binomial
    )),
    results = map(model, ~ tidy(.x, exponentiate = TRUE, conf.int = TRUE))
  ) |>
  unnest(results)

city_or = city_models |>
  filter(term == "victim_sexMale") |>
  select(city_state, estimate, conf.low, conf.high)

city_or
```

Plot showing OR and CI for each city
```{r}
city_or_plot = city_or |>
  arrange(estimate) |>
  mutate(city_state = factor(city_state, levels = city_state))

ggplot(city_or_plot, aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0) +
  labs(
    title = "Adjusted Odds of Solved Homicides for Male vs Female Victims",
    x = "City",
    y = "Adjusted OR for Male vs Female Victimes"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Problem 2

Import data
```{r}
library(p8105.datasets)
data("weather_df")
```

Define linear function
```{r}
boot_stat = function(df) {
  
  fit = lm(tmax ~ tmin + prcp, data = df)
  
  r2 = glance(fit)$r.squared
  
  coefs = tidy(fit) |>
    filter(term %in% c("tmin", "prcp")) 
  
  beta1 = coefs$estimate[coefs$term == "tmin"]
  beta2 = coefs$estimate[coefs$term == "prcp"]
  
  ratio = beta1 / beta2
  
  tibble(r2 = r2, beta_ratio = ratio)
}

```

Run 5,000 bootstrap samples
```{r}
bootstrap_results = tibble(i = 1:5000) |>
  mutate(
    sample = map(i, ~ sample_frac(weather_df, replace = TRUE)),
    stats = map(sample, boot_stat)
  ) |>
  select(stats) |>
  unnest(stats)
```

Plot distribution of estimates
```{r}
bootstrap_results |>
  ggplot(aes(r2)) +
  geom_density() +
  labs(title = "Bootstrap Distribution of R squared")

bootstrap_results |>
  ggplot(aes(beta_ratio)) +
  geom_density() +
  labs(title = "Bootstrap Distribution of Beta 1 / Beta 2")
```

Compute 95% CI
```{r}
bootstrap_results |>
  summarize(
    lower = quantile(r2, 0.025),
    upper = quantile(r2, 0.975)
  )

bootstrap_results |>
  summarize(
    lower = quantile(beta_ratio, 0.025),
    upper = quantile(beta_ratio, 0.975)
  )
```


