---
title: "p8105_hw6_mt3866"
output: github_document
date: "2025-11-17"
---

Load libraries
```{r}
library(tidyverse)
library(broom)
library(modelr)
```

### Problem 1
Import data, and create city_state column, exclude 4 cities, only include White & Black, ensure age is numeric:
```{r}
  homicides = read_csv("homicide-data.csv")|>
  janitor::clean_names()|>
  mutate(city_state = str_c(city, state, sep = ","),
    solved = as.numeric(disposition == "Closed by arrest"),
    victim_age = as.numeric(victim_age)
  ) |>
  filter(!city_state %in% c("Dallas,TX", "Phoenix,AZ", "Kansas City,MO", "Tulsa,AL"))|>
  filter(victim_race %in% c("White", "Black"))
```

Use glm to fit logistic regression, obtain adjusted OR:
```{r}
baltimore_lr = homicides |>
  filter(city_state == "Baltimore,MD")

balt_model = glm(
  solved ~ victim_age + victim_sex + victim_race,
  data = baltimore_lr,
  family = binomial
)

balt_model |>
  broom::tidy(conf.int = TRUE) |>
  filter(term == "victim_sexMale") |>
  mutate(odds_ratio = exp(estimate),
         conf_low = exp(conf.low),
         conf_high = exp(conf.high)) |>
  select(odds_ratio, conf_low, conf_high) |>
  knitr::kable(digits=3)
balt_model
```

Model for all cities, extract adjusted OR:
```{r}
city_models = homicides |>
  group_by(city_state) |>
  nest() |>
  mutate(
    model = map(data, ~ glm(
      solved ~ victim_age + victim_sex + victim_race,
      data = .x,
      family = binomial
    )),
    tidied = map(model, ~ broom::tidy(.x, conf.int = TRUE))
    ) |>
      unnest(tidied) |>
      filter (term == "victim_sexMale")|>
      mutate(odds_ratio = exp(estimate),
            conf_low = exp(conf.low),
            conf_high = exp(conf.high)
            ) |>
      select(city_state, odds_ratio, conf_low, conf_high)

city_models
```

Plot showing OR and CI for each city
```{r}
city_or_plot = city_models |>
  arrange(odds_ratio) |>
  mutate(city_state = factor(city_state, levels = city_state))

ggplot(city_or_plot, aes(x = city_state, y = odds_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf_low, ymax = conf_high), width = 0) +
  labs(
    title = "Adjusted Odds of Solved Homicides for Male vs Female Victims",
    x = "City",
    y = "Adjusted OR for Male vs Female Victims"
  ) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

Comment: The widest 95% CI for the adjusted odds ratio is for the city Albuquerque, NM while the smallest 95% CI for the adjusted odds ratio is for Chicago, IL. The confidence intervals vary widely between the different cities. The city with the highest adjusted odds ratio of solved homicides for males compared to females is in Albuquerque, NM, while the second highest is in Stockton, CA, and the third highest is Fresno, CA. 

### Problem 2

Import data
```{r}
library(p8105.datasets)
data("weather_df")
```

Define linear function
```{r}
boot_stat = function(df) {
  
  fit = lm(tmax ~ tmin + prcp, data = df)
  
  r2 = glance(fit)$r.squared
  
  coefs = tidy(fit) |>
    filter(term %in% c("tmin", "prcp")) 
  
  beta1 = coefs$estimate[coefs$term == "tmin"]
  beta2 = coefs$estimate[coefs$term == "prcp"]
  
  ratio = beta1 / beta2
  
  tibble(r2 = r2, beta_ratio = ratio)
}

```

Run 5,000 bootstrap samples
```{r}
bootstrap_results = tibble(i = 1:5000) |>
  mutate(
    sample = map(i, ~ sample_frac(weather_df, replace = TRUE)),
    stats = map(sample, boot_stat)
  ) |>
  select(stats) |>
  unnest(stats)
```

Plot distribution of estimates
```{r}
bootstrap_results |>
  ggplot(aes(r2)) +
  geom_density() +
  labs(title = "Bootstrap Distribution of R squared")

bootstrap_results |>
  ggplot(aes(beta_ratio)) +
  geom_density() +
  labs(title = "Bootstrap Distribution of Beta 1 / Beta 2")
```
Description: The bootstrap distribution of Beta 1/Beta 2 is left skewed, while the bootstrap distribution of R squared is in the shape of a bell curve.


Identify quantiles, compute 95% CI
```{r}
bootstrap_results |>
  summarize(
    lower = quantile(r2, 0.025),
    upper = quantile(r2, 0.975)
  )

bootstrap_results |>
  summarize(
    lower = quantile(beta_ratio, 0.025),
    upper = quantile(beta_ratio, 0.975)
  )
```

### Problem 3

Import and clean data
```{r}
bw_df = read_csv("birthweight.csv") |>
  mutate(
    babysex = factor(babysex),
    frace   = factor(frace),
    mrace   = factor(mrace),
    malform = factor(malform),
    babysex = factor(babysex, labels = c("male", "female"))
    )|>
  janitor::clean_names()
```

Proposed model
```{r}
proposed_model = lm(
  bwt ~ gaweeks + fincome + smoken + pnumlbw + momage,
  data = bw_df
)
summary(proposed_model)
```

I decided to include these variables in my model because after researching things that may be associated with low birth weight in babies, I found that gestational age in weeks, socioeconomic status, smoking history, preview number of low birthweight babies, and the mom's age are related. Therefore, I included these factors in my model.

Plot of model residuals against fitted values
```{r}
bw_df |>
  add_predictions(proposed_model) |>
  add_residuals(proposed_model) |>
  ggplot(aes(x = pred, y = resid)) +
  geom_point(alpha = 0.4) +
  labs(title = "Residuals vs Fitted Values Proposed Model")
```

Comparison models
```{r}
mod_1 = lm(bwt ~ blength + gaweeks, data = bw_df)

mod_2 = lm(
  bwt ~ bhead * blength * babysex,
  data = bw_df
)
```

Cross validation
```{r}
cv_df = 
  crossv_mc(bw_df, n = 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )
```

Fit all 3
```{r}
cv_results = cv_df |>
  mutate(
    mod_0   = map(train, \(df) lm(bwt ~ gaweeks + fincome + smoken + pnumlbw + momage, data = df)),
    mod_1    = map(train, \(df) lm(bwt ~ blength + gaweeks, data = df)),
    mod_2    = map(train, \(df) lm(bwt ~ bhead * blength * babysex, data = df))
  ) |>
  mutate(
    rmse_0 = map2_dbl(mod_0, test, rmse),
    rmse_1  = map2_dbl(mod_1, test, rmse),
    rmse_2  = map2_dbl(mod_2, test, rmse)
  )
```

Violin plots to compare cross validation prediction error
```{r}
cv_results |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |>
  ggplot(aes(x = model, y = rmse)) +
  geom_violin() +
  labs(title = "Cross-validated RMSE for Birthweight Models")

```

